Pattern matching for $(231,213)$-avoiding permutations

### 2015-01-30 article.tex
correct biblio

	A few particular cases of the pattern containment problem have been attacked successfully.
	The case of increasing patterns is solvable in
	$O(|\pi| \log \log |\sigma|)$ time in the RAM model \cite{Crochemore:Porat:2010},
	improving the previous 30-year bound of $O(|\pi| \log |\sigma|)$.
	(The algorithm also improves on the previous
	$O(|\pi| \log \log |\pi|)$ bound.)
	Furthermore, the patterns $132$, $213$, $231$, $312$ can all be handled in linear time
	by stack sorting algorithms.
	Any pattern of length $4$ can be detected in $O(|\pi| \log |\pi|)$ time
	\cite{Albert:Aldred:Atkinson:Holton:ISAAC:2001}.
	Algorithmic issues for $321$-avoiding patterns containment has been investigated in
	\cite{Guillemot:Vialette:ISAAC:2009}.

	Of particular interest in our context,
	the pattern containment problem is also solvable in
	polynomial time for separable patterns.
	Separable permutations are those permutations that contain neither
	$2413$ nor $3142$, and they are enumerated by the Schröder numbers
	(sequence A006318 in OEIS).
	Separable permutations has gained a lot of interest in the pattern containment
	context during the last years.
	Indeed, the pattern containment problem is solvable $O(kn^4)$ time and
	$O(kn^3)$ space for separable patterns \cite{Ibarra:1997}
	(see also \cite{Bose:Buss:Lubiw:1998}),
	where $k$ is the length of the pattern and $n$ is the length of the
	target permutation.
	Notice that there are numerous characterizations of separable permutations.
	To mention just a few examples,
	they are the permutations whose permutation graphs are cographs
	(\emph{i.e.} $P_4$-free graphs);
	equivalently, a separable permutation is a permutation that can be obtained
	from the trivial permutation $1$ by \emph{direct sums} and \emph{skew sums}.
	To the best of our knowledge,
	separable permutations first arose in the work of
	Avis and Newborn~\cite{Avis:Newborn:1981},
	who showed that they are precisely the permutations which can be sorted by an
	arbitrary number of pop-stacks in series,
	where a pop-stack is a restricted form of stack in which any pop operation
	pops all items at once.
	(Notice that the separable permutations include as a special case the
	stack-sortable permutations, which avoid the pattern $231$.)
	Separable permutations
	have been widely studied in the last decade, both from a combinatorial
	\cite{West:1995} and an algorithmic
	\cite{Ibarra:1997,Bose:Buss:Lubiw:1998,Rossin:Bouvel:06,Bouvel:Rossin:Vialette:CPM:2007} point
	of view.

>>>>

	A few particular cases of the pattern containment problem have been attacked successfully.
	The case of increasing patterns is solvable in
	$O(|\pi| \log \log |\sigma|)$ time in the RAM model \cite{Crochemore:Porat:2010},
	improving the previous 30-year bound of $O(|\pi| \log |\sigma|)$.
	(The algorithm also improves on the previous
	$O(|\pi| \log \log |\pi|)$ bound.)
	Furthermore, the patterns $132$, $213$, $231$, $312$ can all be handled in linear time
	by stack sorting algorithms.
	Any pattern of length $4$ can be detected in $O(|\pi| \log |\pi|)$ time
	\cite{Albert:Aldred:Atkinson:Holton:ISAAC:2001}.
	Algorithmic issues for $321$-avoiding patterns containment has been investigated in
	\cite{Guillemot:Vialette:ISAAC:2009}.
	The pattern containment problem is also solvable in
	polynomial-time for separable patterns \cite{Ibarra:1997}
	(see also \cite{Bose:Buss:Lubiw:1998}).
	Separable permutations are those permutations that contain neither
	$2413$ nor $3142$, and they are enumerated by the Schröder numbers
	(sequence A006318 in OEIS).
	To the best of our knowledge,
	separable permutations first arose in the work of
	Avis and Newborn~\cite{Avis:Newborn:1981},
	who showed that they are precisely the permutations which can be sorted by an
	arbitrary number of pop-stacks in series,
	where a pop-stack is a restricted form of stack in which any pop operation
	pops all items at once.
	(Notice that the separable permutations include as a special case the
	stack-sortable permutations, which avoid the pattern $231$.)
